---
title: "Initial Korean covid 19"
author: "Ashley Saunders"
date: "9/26/24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

At 55%, the majority of those alive were female, with males at 45%. However, out of those who died, 65% were male and 45% were female. This shows a significant difference between those who died and didn't, majority being male. 

## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The data shows evidence that the majority of the alive population were female, with a ratio of 56:44. Furthermore, the dead population was majority male, with a ratio of 36:64. 3% of males were dead and 1.4% of females were also dead.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

According to the Null results, there should be an equal distribution of dead and alive male and female individuals. The observed counts, however, revealed that the proportion of dead women was almost twice as high as that of men. Men are more likely than women to die, as seen by the 2:3 ratio of female to male deaths.

## Conclusions

Despite the fact that women make up the bulk of the population, the facts and graphs shown indicate a correlation between sex and death, with men being more than twice as likely to die as women.

